Analysis of split plot design using MCMCglmm

### 2017-08-01 Analysis.Rmd
Clarify analysis further
# Main changes

- Reorder base factors for analysis to make interpretation easier

- Add more detail on how the MCMCglmm models are set up and should be
interpreted

- Add more detail interpreting the model outputs

# Minor changes

- Remove probability distribution checks

- Update README.md

- Add links to more resources

## Notes on this analysis before starting

@@

#### How to interpret these analyses

@@

@@

Before the analysis, note that due to changes between the years, the analysis
must be carried out on each year separately. The 2015 data and 2016 data cannot
be combined due to changes in inoculation methods; there are other changes too,
but the main one is this. Therefore, the analyses will be conducted separately for
leaf severity and Leaf severity for 2015 dry season and 2016 dry season.
Comparisons will only be observational and cannot be statistically compared.

#### Iterations

@@

First, separate the TRT column into the two treatments for analysis

@@

```{r create_analysis_df}
AUDPS_15 <- as.data.frame(AUDPS[AUDPS$YEAR == 2015, ])
AUDPS_15 <- droplevels(AUDPS_15)

AUDPS_16 <- as.data.frame(AUDPS[AUDPS$YEAR == 2016, ])
AUDPS_16 <- droplevels(AUDPS_16)
```

Now that the `AUDPS_15` and `AUDPS_16` `data.frames` exist, we can start the analysis.

@@

### Tiller Sheath Blight Incidence and Severity

In 2015 and 2016 the highest N treatment was significant in both the tiller sheath blight incidence and severity. In 2016 the flooding treatment was significant for tiller sheath blight severity.

In neither experiment was the leaf sheath blight affected by nitrogen rates or irrigation regime.

### Interactions of Nitrogen and Irrigation Regime

In 2015 no significant itneraction was observed. In 2016 a significant interaction between flooding and nitrogen rate is observed for both the tiller and leaf severity ratings.

@@

None of the diagnostic plots show any signs of autocorrelation or any other obvious patterns.

@@

The random effects are all fairly equally distributed except for 2016 tiller sheath blight severity where water management:replicate has a larger effect. This is not surprising given that the plot and replicate sizes were different in 2016 due to the use of two fields in the IRRI experiment station, see the [2016 plot layout plan](https://github.com/adamhsparks/AWD_ShB_Interaction/blob/master/doc/Metadata/2016%20AWD%20ShB%20Metadata/ExptLayoutAWD2016.pdf).

>>>>

## Notes on This Analysis Before Starting

If you have not already looked at the
[01-Probability_distribution_checks](01-Probability_distribution_checks.md),
please see that first to understand why MCMCglmm was chosen for this analysis
due to the nature of the data.

@@

#### Interpreting these analyses

@@

I found this CrossValidated topic to be useful in interpreting the outputs from
MCMCglmm as it explains the interpretation of the effects. While it is about
output from a `lmer`, the output is read in the same way, [How to interpret 2-way and 3-way interaction in lmer?](https://stats.stackexchange.com/questions/87412/how-to-interpret-2-way-and-3-way-interaction-in-lmer#87415).

The base levels for this analysis are:

- 2015 `NRTE:0`

- 2015 `WMGT:FLD`

- 2016 `NRTE:60`

- 2016 `WMGT:FLD`

@@

Before the analysis, note that due to changes between the years, the analysis
must be carried out on each year separately. The 2015 data and 2016 data cannot
be combined due to changes in inoculation methods; there are other changes too,
but the main one is this. Therefore, the analyses will be conducted separately
such that comparisons will only be observational and cannot be statistically
compared.

#### MCMCglmm model fitting notes

@@

First, separate the TRT column into the two treatments for analysis.

@@

```{r create_analysis_df}
AUDPS_2015 <- as.data.frame(AUDPS[AUDPS$YEAR == 2015, ])
AUDPS_2015 <- droplevels(AUDPS_2015)

# relevel factors for easier interpretation of  analysis
AUDPS_2015 <- within(AUDPS_2015, NRTE <- relevel(NRTE, ref = "N0"))
AUDPS_2015 <- within(AUDPS_2015, WMGT <- relevel(WMGT, ref = "FLD"))

AUDPS_2016 <- as.data.frame(AUDPS[AUDPS$YEAR == 2016, ])
AUDPS_2016 <- droplevels(AUDPS_2016)

# relevel factors for easier interpretation of  analysis
AUDPS_2016 <- within(AUDPS_2016, NRTE <- relevel(NRTE, ref = "N60"))
AUDPS_2016 <- within(AUDPS_2016, WMGT <- relevel(WMGT, ref = "FLD"))
```

Now that the `AUDPS_2015` and `AUDPS_2016` `data.frames` exist, we can start the analysis.

@@

@@

None of the diagnostic plots show any signs of auto-correlation or any other obvious patterns.

@@

### Tiller Sheath Blight Incidence

In 2015 the highest N rate, N120, was significantly different from the base, N0 rate.

In 2016 both the AWD and highest N rate, N180, were significantly different from the FLD and N60 treatments.

No interactions were found to be significant in the tiller sheath blight incidence.

## Tiller Sheath Blight Severity

In 2015 both the N100 and N120 rates were significantly different from the N0 treatment. The interaction of the N100 rate with AWD was also significantly different than N0 and FLD treatments.

In 2016 the N180 was significantly different from the N60 rate and the interaction of the N180 and AWD was significantly different than the N60 and FLD interaction.

In 2015 the N120 was significantly different than the N0 treatment. The interactions of both the N100 and N120 with AWD were significantly different than the N0 FLD interaction.

In 2016 the N180 was significantly different from the N60 treatment. The AWD was significantly different from the FLD and the interaction of N180:AWD was significantly different from the N60:FLD treatments interaction.
